---
title: "Part 6. Lists and RMarkdown"
linktitle: "Part 6. Lists and RMarkdown"
date: "2021-02-10"
class_date: "2021-02-10"
output:
  blogdown::html_page:
    toc: true
menu:
  class:
    parent: Class sessions
    weight: 6
type: docs
weight: 6
# pdf: 
# thumb: 
editor_options: 
  chunk_output_type: console
---

## Class Video

<iframe width="560" height="315" src="https://www.youtube.com/embed/7c3xGQBmjIs" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

## Slides

Open the slides in a separate window: <a href="https://sph-r-programming.netlify.com/slides/06_intro_to_rmarkdown#1" target="_blank">https://sph-r-programming.netlify.com/slides/06_intro_to_rmarkdown#1</a>

```{r echo=FALSE}
knitr::include_url('https://sph-r-programming.netlify.com/slides/06_intro_to_rmarkdown#1')
```

## Midterm assigned (Due February 12 @ 11:55 PM)

RStudio Project is now available. Please come to office hours or stay for class to discuss your data set.

### Post-Class

Please fill out the following survey and we will discuss the results during the next lecture. All responses will be anonymous.

- Clearest Point: What was the most clear part of the lecture?
- Muddiest Point: What was the most unclear part of the lecture to you?
- Anything Else: Is there something you'd like me to know?

http://bit.ly/sph504_survey

## Muddiest Points

> Lists are a little confusing. I think the homework will help.

> Lists are still very new, so I'm planning to go through the part 6 file again.

Yes, lists are weird. But after part 7, hopefully you'll see they're very useful.

> Still trying to wrap my head around [[]] vs [].

I will say that almost 99% of the time, you should be using double brackets `[[]]`, because you want what's in the list slot. 


```{r}
my_list <- list(cat_names = c("Morris", "Julia"), 
                hedgehog_names = "Spiny", 
                dog_names = c("Rover", "Spot"))
```

```{r}
df <- my_list[["cat_names"]]
class(df)
```

Again, using `[]` returns a list of length 1, which is usually not what you want.

```{r}
single_list <-  my_list["cat_names"]
class(single_list)
single_list
```

To actually access the `character` vector in `single_list`, we need to access it out of the list structure with `[[]]`.

```{r}
single_list[["cat_names"]]
```

> Posting r html file on the website

> Websites

That was a really quick demo and I will go over it more in detail when we get to working with websites.

If you want a preview, you can check out [Sharing Online on Short Notice](https://rstudio.com/resources/webinars/sharing-on-short-notice-how-to-get-your-materials-online-with-r-markdown/.)
