---
title: 'scales::breaks_pretty()'
author: "Sarah Gothard"
date: "`r Sys.Date()`"
menu:
  functions:
    weight: 7
type: docs
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Sarah Gothard


In this document, I will introduce the breaks_pretty() function and show what it's for.

```{r}
library(tidyverse)
library(palmerpenguins)
library(scales)
data(penguins)

```

The "breaks_pretty" function is used to find best fit "breaks" or intervals in a data set. 
```{r}

#creating simple breaks in a month
one_month <- as.POSIXct(c("2020-05-01", "2020-06-01"))
demo_datetime(one_month, breaks = breaks_pretty(2))
demo_datetime(one_month, breaks = breaks_pretty(4))

```
```{r}
#more complex breaks and/or data sets will sometimes lead to greater or fewer intervals than desgnated by your "n" because the breaks_pretty package intuitively finds breaks near your desired number of intervals that fits the data assigned best.

demo_datetime(one_month, breaks = breaks_pretty(12))
#output has more than 12 breaks
```

Funciton Discussion and Use

Breaks_pretty is found in the "scales" package and takes advantage of an existing r package/algorithm, "pretty()". In practice, breaks_pretty() is most commonly used to designate the tics/intervals of dates/times/continuous variables in ggplot visualizations. 
```{r}
#when using breaks_pretty() add you can either designate your argument's n explicity, (n = 6) or it defaults to that as the first argument and pass in just the number of breaks you'd like to see (6).

ggplot(penguins,
       aes(x = bill_length_mm,
           y = body_mass_g)) +
  geom_point() +
  geom_smooth() +
  scale_x_continuous(name = "Bill Length (mm)",
                     breaks = breaks_pretty(6)) +     # breaks_pretty
  scale_y_continuous(name = "Body Mass (g)",
                     breaks = breaks_pretty(n = 8)) +
  theme_light()

```

Is it helpful?

Breaks_pretty is very useful as it simplifies how we desginate intervals (tics) in our ggplots. Instead of hardcoding the intervals we want to see, breaks_pretty intuitively designates tics close to our desired widths.  
