---
title: 'dplyr::slice_max()'
author: "Bradley Hopkins"
date: "`r Sys.Date()`"
menu:
  functions:
    weight: 3
type: docs
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Bradley Hopkins

In this document, I will introduce the `slice_max()` function and show what it's for.

```{r}
#load tidyverse up
library(tidyverse)
library(dplyr)
```

## What is it for?

`slice_max()` is a sub-function of `slice()`, the function we discussed last week. Like `slice()`, `slice_max()` also makes it easier to visualize a data frame by providing a look at what is contained inside. What makes `slice_max()` unique is that it allows us to specify a variable we want to sort by and then returns the maximum values for that variable in the number of rows we specify. The difference between `slice_()` and `slice_max()` is demonstrated below.

```{r}
mtcars %>% slice(n = 1:5)
```

```{r}
mtcars %>% slice_max(order_by = mpg, n = 5)
```

Take note of how the output for `slice_max()` has returned the rows of `mtcars` with the highest values for the variable `mpg` in descending order.

## Is it helpful?

This function is similar to its parent function `slice()`, but fills a slightly different niche. As with `slice()`, `slice_max()` is most useful because it can be used with `%>%`. This makes it slightly easier to pare down and organize a set of data compared to entering a `[row, column]` reference with the data frame. This function can also be combined with other methods of sorting data to produce more interesting results. In the example below, we filter `mtcars` down to only 6-cylinder cars first and then pipe into `slice_max()` to produce a list of 6-cylinder cars with the highest mpg:

```{r}
mtcars %>% filter(cyl == 6) %>% slice_max(order_by = mpg, n = 5)
```

This function is useful for visualizing data, providing users with a tool that can help users better understand the hierarchy of certain variables. It may even provide some utility in organizing and sorting data if used as part of a more complex pipeline of functions.
