---
title: 'stringr::str_wrap()'
author: "Sam Callis"
date: "`r Sys.Date()`"
menu:
  functions:
    weight: 8
type: docs
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Sam Callis

In this document, I will introduce the `str_wrap()` function and show what it's for.

```{r}
#load tidyverse up
library(tidyverse)

#example dataset
string1 <- "This is a string"
string2 <- "Most programming languages have a data type called a string, which is used for data values that are made up of ordered sequences of characters, such as 'hello world'. A string can contain any sequence of characters, visible or invisible, and characters may be repeated."
```

## What is it for?

The str_wrap() function gives us a way to wrap strings into paragraphs. The formatting of the paragraphs can be adjusted by width and can include an indent or exdent. str_wrap() is typically combined with the cat() (concatenate) function.

```{r}
#without cat() function
str_wrap(string2, width = 80)

#with cat() function
cat(str_wrap(string2, width = 80), "\n")
cat(str_wrap(string2, width = 40, indent = 2),"\n")
cat(str_wrap(string1, width = 1, exdent = 2), "\n")
```

## Is it helpful?

I have not personally worked with many datasets that includes strings, but I would imagine str_wrap() would be very helpful for managing them. Especially if the strings are long.
